### 2021-05-31 article/article.tex
Article draft 4

\section*{(1) Overview}

@@

\section*{Title}
spinsim: a GPU optimised simulator of spin half and spin one quantum systems

\section*{Paper Authors}
1. Tritt, Alex;\\
2. Morris, Joshua;\\
3. Hochstetter, Joel;\\
4. Anderson, R. P.;\\
5. Saunderson, James;\\
6. Turner, L. D.;\\

\section*{Paper Author Roles and Affiliations}
1. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Primary author of the released packages.\\
2. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Present address: Faculty of Physics, University of Vienna, 1010 Vienna, Austria.\\
	Author of first version of code.\\
3. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Present address: School of Physics, University of Sydney, NSW 2006, Australia.\\
	Optimization and extension to spin one of first version of code.\\
4. School of Molecular Sciences, La Trobe University, PO box 199, Bendigo, Victoria 3552, Australia.\\
	Original conception of first version of code.\\
5. Department of Electrical and Computer Systems Engineering, Monash University, Victoria 3800, Australia.\\
	Advice on numerical analysis.\\
6. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Original conception of released version of algorithm.

\section*{Abstract}
	\texttt{spinsim} is a \emph{python} package that simulates spin half and spin one quantum mechanical systems following a time dependent Shroedinger equation. It makes use of \texttt{numba.cuda} \cite{lam_numba_2015}, which is an \emph{LLVM} (Low Level Virtual Machine) \cite{lattner_llvm_2004} compiler, and other optimisations, to allow for fast and accurate evaluation on \emph{Nvidia Cuda} \cite{nickolls_scalable_2008} compatible systems using GPU parallelisation. \texttt{spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}. The initial use case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin half or spin one quantum systems with time dependent Hamiltonians that cannot be solved analytically. These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin one systems of nitrogen vacancy centres (NVCs) and BECs.

\section*{Keywords}
Time dependent Schroedinger equation; Spin one; Spin half; Integrator; GPU; Solver; python; numba;

\section*{Introduction}
\subsection*{Motivation}
	Ultracold rubidium atoms have proven their effectiveness in state of the art technologies in quantum sensing \cite{degen_quantum_2017}, the use of quantum mechanics to make precise measurements of small signals. The rotation of these atoms can be modelled as quantum spin systems, which is the quantum mechanical model for objects with angular momentum. The simplest spin system, spin half (ie spin quantum number of \(\frac12\), also referred to as a qubit), is quantised into just two quantum spin levels, and this describes the motion of some fundamental particles such as electrons. However, systems more practical for sensing, such as ultracold rubidium atoms, are more accurately described as a spin one quantum system (ie spin quantum number of \(1\), also referred to as a quitrit), which is quantised into three quantum spin levels.

@@

\section*{Implementation and architecture}
	\subsection*{Mathematical methods}
		\subsubsection*{Background}
			In general, \texttt{spinsim} solves the Schroedinger equation,

			\begin{align}
				\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) &= -iH(t)\psi(t),
			\end{align}

			where \(i^2 = -1\), the quantum state \(\psi(t) \in \mathbb{C}^N\) a time dependent, \(N\) dimensional, unit complex vector, and the Hamiltonian \(H(t) \in \mathbb{C}^{N \times N}\) is a time dependent \(N \times N\) complex Hermitian matrix. Here \(N\) is the number spin levels in the quantum system, so spin half is the \(N = 2\) case, and spin one refers to the \(N = 3\) case. Rather than being represented by standard coordinates in \(\mathbb{C}^{N \times N}\), in \texttt{spinsim} the Hamiltonian \(H(t)\) is instead represented with respect to a choice of basis for the corresponding Lie Algebra, \(\mathfrak{su}(N)\). For example, when set to spin half mode, the \texttt{spinsim} package solves the time dependent Schroedinger equations of the form

@@

			\begin{align}
				J_x &= \frac12\begin{pmatrix}
					0 & 1 \\
					1 & 0
				\end{pmatrix},
				&J_y &= \frac12\begin{pmatrix}
					0 & -i \\
					i &  0
				\end{pmatrix},
				&\textrm{and }J_z &= \frac12\begin{pmatrix}
					1 &  0 \\
					0 & -1
				\end{pmatrix}.
			\end{align}

			The energy source \(f\) that represents the time dependent Hamiltonian \(H(t)\), is the collection of energy functions \(f_x(t), f_y(t), f_z(t)\), with \(t\) in units of s and \(f\) in units of Hz that control the dynamics of the system. The user must define a method that returns a sample of these source functions when a sampling time is input. In physical terms, these functions could represent the \(x,y,z\) components of a magnetic field applied to a magnetically sensitive.

@@

			\begin{align}
				J_x &= \frac{1}{\sqrt{2}}\begin{pmatrix}
					0 & 1 & 0 \\
					1 & 0 & 1 \\
					0 & 1 & 0
				\end{pmatrix},&
				J_y &= \frac{1}{\sqrt{2}}\begin{pmatrix}
					0 & -i &  0 \\
					i &  0 & -i \\
					0 &  i &  0
				\end{pmatrix},\nonumber\\
				J_z &= \begin{pmatrix}
					1 & 0 &  0 \\
					0 & 0 &  0 \\
					0 & 0 & -1
				\end{pmatrix},&
				\textrm{and }Q &= \frac{1}{3}\begin{pmatrix}
					1 &  0 & 0 \\
					0 & -2 & 0 \\
					0 &  0 & 1
				\end{pmatrix}.
			\end{align}

@@

		\subsubsection*{Parallelisation}
			Given that \(\psi(t)\) is a unit vector, it is possible to write \(\psi(t)\) in terms of a unitary transformation \(U(t, t_0)\) of the state \(\psi(t_0)\), for any time \(t_0\). In other words, for any \(t_0,t \in \mathbb{R}\), there is a unitary transformation \(U(t, t_0)\) such that

@@

			This means that a time series for the state of the system can be evaluated by evaluating the time evolution operator between each of the sample times. Consider quantising time with
			\begin{align}
				t_k &= t_0 + \mathrm{D}t\cdot k,
			\end{align}

@@

		\subsubsection*{Rotating frame}
			If the rotating frame option is selected, the \(U_k\) are first calculated within a rotating frame of reference as \(U^r_k\), which, in some situations, reduces the size of the source functions used in the calculation, increasing accuracy. The rotation speed of the rotating frame is calculated locally for each parallel time step \(U_k\), and only for rotations around the \(z\) axis. The rotating from source functions \(f^r_x, f^r_y, f^r_z,\) and \(f^r_q\) are related to the source function from the user input via

@@

			Where \(f_r = f_z(t_k + \frac12\mathrm{D}t)\) is sampled the midpoint value of the source over the interval \([t_{k - 1}, t_k]\). This, assuming that a midpoint sample is representative of an average value over the time interval, decreases the magnitude of \(f_z(t)\), while leaving the other source components at an equivalent magnitude. The rotation is then applied to obtain the lab frame time evolution operator \(U_k\) via

@@

		\subsubsection*{Magnus based integration method}
			The integration method used in \texttt{spinsim} is the commutator free 4 (CF4) method from Auer et al \cite{auer_magnus_2018}, which based on the Magnus expansion. Each of the \(U_k\) are split into products of time evolution operators between times separated by a smaller time step, that is,

			The CF4 method is used to calculate each individual \(u^k_l\). Let the fine sample time be given by \(t_f = l\mathrm{d}t + t_k\). Then as part of the CF4 method, the source functions are sampled at particular times based on the second order Gauss-Legendre quadrature, given by

			Now, let
			\begin{align}
				f(t_1) &= (f_x(t_1), f_y(t_1), f_z(t_1), f_q(t_1))\textrm{, and}
				f(t_2) &= (f_x(t_2), f_y(t_2), f_z(t_2), f_q(t_2)).
			\end{align}

			The integration time evolution operator can then be calculated using

			\begin{align}
				g_1 =& (g_{1,x}, g_{1,y}, g_{1,z}, g_{1,q})\\
				=& 2 \pi \mathrm{d}t \left(\frac{3 + 2 \sqrt{3}}{12} f(t_1) + \frac{3 - 2 \sqrt{3}}{12} f(t_2)\right)\textrm{, and}\\
				g_2 =& (g_{2,x}, g_{2,y}, g_{2,z}, g_{2,q})\\
				=& 2 \pi \mathrm{d}t \left(\frac{3 - 2 \sqrt{3}}{12} f(t_1) + \frac{3 + 2 \sqrt{3}}{12} f(t_2)\right)\textrm{, so}\\
				u =& \exp(-i \left( g_{2,x} J_x + g_{2,y} J_y + g_{2,z} J_z + g_{2,q} Q\right))\\
				&\cdot\exp(-i \left( g_{1,x} J_x + g_{1,y} J_y + g_{1,z} J_z + g_{1,q} Q\right)).
			\end{align}

		\subsubsection*{Exponentiator}
			For all exponentiation, the exponentiator computes the matrix exponential

			\begin{align}
				E(g) &= E(g_x, g_y, g_z, g_q)\\
				&= \exp(-i (g_x J_x + g_y J_y + g_z J_z + g_q Q)), \textrm{ with}
			\end{align}

			For spin half, the default exponentiator is in an analytic form. For spin one, an exponentiator based on the Lie Trotter product formula \cite{moler_nineteen_2003} is used instead. Explicitly, this formula is

			Where \(A,B\in\mathbb{C}^{N\times N}\) are matrix operators, and \(n\in\mathbb{N}\). The equality holds for without the limit for all \(n\) if the operators commute, like it does for real and complex numbers. This tells us that we can approximate a matrix exponential of a linear combination of spin operators by reducing the size of the coefficients of these operators, taking the analytically known exponentials of each of the operators separately, multiplying each individual result together, and then raising the combination \(T\) to the power of the original reduction. For our spin one case, the exponential \(E(g)\) can be approximated as, for large \(2^\tau\),

            % \begin{align}
            %     E(g) &= \exp(-ig_x J_x - ig_y J_y - ig_z J_z - ig_q Q)\\
            %     &= \exp(2^{-\tau}(-ig_x J_x - ig_y J_y - ig_z J_z - ig_q Q))^{2^\tau}\\
            %     &\approx (\exp(-i(2^{-\tau} g_x) J_x) \exp(-i(2^{-\tau} g_y) J_y) \exp(-i(2^{-\tau} g_z J_z + (2^{-\tau} g_q) Q)))^{2^\tau}\\
            %     &= \begin{pmatrix}
            %         \frac{e^{-i\left(Z + \frac{P}{3}\right)}(c_X + c_Y - i s_Xs_Y)}{2} & \frac{e^{i\frac{2P}{3}} (-s_Y -i c_Y s_X)}{\sqrt{2}} & \frac{e^{-i\left(-Z + \frac{P}{3}\right)}(c_X - c_Y + i s_Xs_Y)}{2} \\
            %         \frac{e^{-i\left(Z + \frac{P}{3}\right)} (-i s_X + c_X s_Y)}{\sqrt{2}} & e^{i\frac{2P}{3}} c_X c_Y & \frac{e^{-i(Z - \frac{P}{3})} (-i s_X - c_X s_Y)}{\sqrt{2}} \\
            %         \frac{e^{-i\left(Z + \frac{P}{3}\right)}(c_X - c_Y - i s_Xs_Y)}{2} & \frac{e^{i\frac{2P}{3}} (s_Y -i c_Y s_X)}{\sqrt{2}} & \frac{e^{-i\left(-Z + \frac{P}{3}\right)}(c_X + c_Y + i s_Xs_Y)}{2}
            %     \end{pmatrix}^{2^\tau}
            % \end{align}

			\begin{align}
                E(g) =& \exp\left(-ig_x J_x - ig_y J_y - ig_z J_z - ig_q Q\right)\\
                =& \exp\left(2^{-\tau}\left(-ig_x J_x - ig_y J_y - ig_z J_z - ig_q Q\right)\right)^{2^\tau}\\
                \approx& \biggl(\exp\left(-i\left(2^{-\tau} \frac12g_z J_z + \left(2^{-\tau}\frac12g_q\right) Q\right)\right)\nonumber\\
				&\cdot\exp\left(-i\left(2^{-\tau} g_\phi J_\phi\right)\right)\nonumber\\
				&\cdot\exp\left(-i\left(2^{-\tau} \frac12g_z J_z + \left(2^{-\tau} \frac12g_q\right) Q\right)\right)\biggr)^{2^\tau}\\
                =& \begin{pmatrix}
                    \left(\frac{\Gamma}{PZ}\right)^2 & \frac{SP}{Z\Phi} & -\left(\frac{\Sigma P}{\Phi}\right)^2 \\
					\frac{SP\Phi}{Z} & CP^4 & \frac{SPZ}{\Phi} \\
					-\left(\frac{\Sigma \Phi}{P}\right)^2 & \frac{SPZ\Phi}{1} & \left(\frac{\Gamma Z}{P}\right)^2
                \end{pmatrix}^{2^\tau}\\
				=& T^{2^\tau}.
            \end{align}

            \begin{align}
				g_\phi &= \sqrt{g_x^2 + g_y^2}&&\nonumber\\
				J_\phi &= \frac{g_x}{g_\phi}J_x + \frac{g_y}{g_\phi}J_y&&\nonumber\\
				\Gamma &= \cos\left(2^{-\tau} \frac{g_\phi}{2}\right) & \Phi &= e^{i2^{-\tau}g_\phi}\nonumber\\
				\Sigma &= \sin\left(2^{-\tau} \frac{g_\phi}{2}\right) & Z &= e^{i2^{-\tau}\frac{g_z}{2}}\nonumber\\
				C &= \cos(2^{-\tau} g_\phi) & P &= e^{i2^{-\tau}\frac{g_q}{6}}\nonumber\\
				S &= \frac{-i}{\sqrt2}\sin(2^{-\tau} g_\phi)&&
            \end{align}

			Note that, the methods for both spin half and spin one use analytic forms of exponentials to construct the result, meaning that all calculated time evolution operators are unitary. This guarantees that the results of \texttt{spinsim} maintain unitary.

	\subsection*{Software architecture}
		\subsubsection*{Integrator architecture}
			The integrator in the \texttt{spinsim} package calls a \texttt{numba.cuda.jit()}ed kernel to be run on a \emph{Cuda} capable \emph{Nvidia} GPU in parallel, with a different thread being allocated to each of the \(U_k\). This returns when each of the \(U_k\) have been evaluated.

			The thread starts by calculating \(t_k\) and, if the rotating frame is being used, \(f_r\). The latter is done by sampling a (\texttt{numba.cuda.jit()}ed version of a) user provided \emph{python} function \(f\) describing how to sample the source Hamiltonian. The code then loops over each integration time step \(\mathrm{d}t\) to calculate the integration time evolution operators \(u^k_l\).

@@

			where \(j \in \{\frac12, 1\}\) is the spin quantum number of the system. This simulation of the quantum sensing experiment involves continuously driving transitions in the system, while exposing it to a short pulsed signal that the system should be able to sense. It runs over a duration of 1ms.

>>>>

\section{(1) Overview}

@@

\section{Title}
spinsim: a GPU optimised simulator of spin half and spin one quantum systems

\section{Paper Authors}
1. Tritt, Alex;\\
2. Morris, Joshua;\\
3. Hochstetter, Joel;\\
4. Anderson, R. P.;\\
5. Saunderson, James;\\
6. Turner, L. D.;\\

\section{Paper Author Roles and Affiliations}
1. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Primary author of the released packages.\\
2. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Present address: Faculty of Physics, University of Vienna, 1010 Vienna, Austria.\\
	Author of first version of code.\\
3. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Present address: School of Physics, University of Sydney, NSW 2006, Australia.\\
	Optimization and extension to spin one of first version of code.\\
4. School of Molecular Sciences, La Trobe University, PO box 199, Bendigo, Victoria 3552, Australia.\\
	Original conception of first version of code.\\
5. Department of Electrical and Computer Systems Engineering, Monash University, Victoria 3800, Australia.\\
	Advice on numerical analysis.\\
6. School of Physics \& Astronomy, Monash University, Victoria 3800, Australia.\\
	Original conception of released version of algorithm.

\section{Abstract}
	\texttt{spinsim} is a \emph{python} package that simulates spin half and spin one quantum mechanical systems following a time dependent Shroedinger equation. It makes use of \texttt{numba.cuda} \cite{lam_numba_2015}, which is an \emph{LLVM} (Low Level Virtual Machine) \cite{lattner_llvm_2004} compiler, and other optimisations, to allow for fast and accurate evaluation on \emph{Nvidia Cuda} \cite{nickolls_scalable_2008} compatible systems using GPU parallelisation. \texttt{spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}. The initial use case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin half or spin one quantum systems with time dependent Hamiltonians that cannot be solved analytically. These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin one systems of nitrogen vacancy centres (NVCs) and BECs.

\section{Keywords}
Time dependent Schroedinger equation; Spin one; Spin half; Integrator; GPU; Solver; python; numba;

\section{Introduction}
\subsection{Motivation}
	Ultracold rubidium atoms have proven their effectiveness in state of the art technologies in quantum sensing \cite{degen_quantum_2017}, the use of quantum mechanics to make precise measurements of small signals. The rotation of these atoms can be modelled as quantum spin systems, which is the quantum mechanical model for objects with angular momentum. The simplest spin system, spin half (ie spin quantum number of \(\frac12\), also referred to as a qubit), is quantised into just two quantum spin levels, and this describes the motion of some fundamental particles such as electrons. However, systems more practical for sensing, such as ultracold rubidium atoms, are more accurately described as a spin one quantum system (ie spin quantum number of \(1\), also referred to as a quitrit), which is quantised into three quantum spin levels.

@@

\section{Implementation and architecture}
	\subsection{Mathematical methods}
		\subsubsection{Quantum mechanics background}
			In general, \texttt{spinsim} solves the Schroedinger equation,

			\begin{align}
				\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) &= -i2\pi H(t)\psi(t),\label{eq:schroedinger}
			\end{align}

			where \(i^2 = -1\), the quantum state \(\psi(t) \in \mathbb{C}^N\) a time dependent, \(N\) dimensional, unit complex vector, and the Hamiltonian \(H(t) \in \mathbb{C}^{N \times N}\) is a time dependent \(N \times N\) complex Hermitian matrix. Here \(N\) is the number spin levels in the quantum system, so spin half is the \(N = 2\) case, and spin one refers to the \(N = 3\) case. Furthermore, we have chosen for the Hamiltonian \(H(t)\) to have the physical dimension of frequency, rather than the more typical dimension of energy. To elaborate, we represent energies \(E\) of the system as their equivalent frequencies \(f\) from the Planck relation

			\begin{align}
				E &= \hbar \omega\\
				&= 2\pi\hbar f,
			\end{align}

			where \(\hbar\) is (reduced) Planck's constant (or equivalently, setting 
			\(\hbar = 1\)). In addition, rather than being represented by standard coordinates in \(\mathbb{C}^{N \times N}\), in \texttt{spinsim} the Hamiltonian \(H(t)\) is instead represented with respect to a choice of basis for the corresponding Lie Algebra, \(\mathfrak{su}(N)\). For example, when set to spin half mode, the \texttt{spinsim} package solves the time dependent Schroedinger equation of the form

@@

			\begin{align}
				J_x &= \frac12\begin{pmatrix}
					0 & 1 \\
					1 & 0
				\end{pmatrix},
				&J_y &= \frac12\begin{pmatrix}
					0 & -i \\
					i &  0
				\end{pmatrix},
				&\textrm{and }J_z &= \frac12\begin{pmatrix}
					1 &  0 \\
					0 & -1
				\end{pmatrix}.\label{eq:spin_half_operators}
			\end{align}

			The fields \(f\) that represents the time dependent Hamiltonian \(H(t)\), are the collection of energy functions \(f_x(t), f_y(t), f_z(t)\), with time \(t\) and \(f\) with physical dimension of frequency that control the dynamics of the system. The user must define a method that returns a sample of these field functions when a sampling time is input. In physical terms, these functions could represent the \(x,y,z\) components of a magnetic field applied to a magnetically spin one BEC or NVC.

@@

			\begin{align}
				J_x &= \frac{1}{\sqrt{2}}\begin{pmatrix}
					0 & 1 & 0 \\
					1 & 0 & 1 \\
					0 & 1 & 0
				\end{pmatrix},&
				J_y &= \frac{1}{\sqrt{2}}\begin{pmatrix}
					0 & -i &  0 \\
					i &  0 & -i \\
					0 &  i &  0
				\end{pmatrix},\nonumber\\
				J_z &= \begin{pmatrix}
					1 & 0 &  0 \\
					0 & 0 &  0 \\
					0 & 0 & -1
				\end{pmatrix},&
				\textrm{and }Q &= \frac{1}{3}\begin{pmatrix}
					1 &  0 & 0 \\
					0 & -2 & 0 \\
					0 &  0 & 1
				\end{pmatrix}.\label{eq:spin_one_operators}
			\end{align}

@@

		\subsubsection{Unitary time evolution and the Magnus expansion}
			Given that \(\psi(t)\) is a unit vector, it is possible to write \(\psi(t)\) in terms of a unitary transformation \(U(t, t_0)\) of the state \(\psi(t_0)\), for any time \(t_0\). In other words, for any \(t_0,t \in \mathbb{R}\), there is a unitary transformation \(U(t, t_0)\) such that

@@

			From Equation \eqref{eq:schroedinger}, \(U(t, t_0)\) also follows the Schroedinger equation via

			\begin{align}
				\frac{\mathrm{d}}{\mathrm{d}t}U(t, t_0)\psi(t_0) &= -i2\pi H(t)U(t, t_0)\psi(t_0)\textrm{ for any }\psi(t_0)\in\mathbb{C}^N,\textrm{ so}\\
				\frac{\mathrm{d}}{\mathrm{d}t}U(t, t_0) &= -i2\pi H(t)U(t, t_0).\label{eq:unitary_schroedinger}
			\end{align}

			Thus, to solve Equation \eqref{eq:schroedinger} for \(\psi(t)\) given a particular \(\psi(t_0)\), one only needs to solve Equation \eqref{eq:unitary_schroedinger} for \(U(t, t_0)\). Note that because \(U(t, t_0)\) is unitary, the calculated value of \(\psi(t)\) using this method will not change in magnitude, conserving probability.

			Also since \(U(t, t_0)\) is unitary (and continuous in time \(t\)), it can be written as a matrix exponential of a skew Hermitian matrix. For example, if the Hamiltonian \(H(t)\) is constant in time (and a real linear combination of Hermitian matrices defined by Equations \eqref{eq:spin_half_operators} or \eqref{eq:spin_one_operators}), then the solution would be

			\begin{align}
				U(t, t_0) &= \exp(-i2\pi H(t - t_0)).
			\end{align}

			The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series \cite{blanes_magnus_2009}, where

			\begin{align}
				U(t, t_0) &= \exp(-i2\pi G(t - t_0))\\
				&= \exp(-i2\pi \sum_{m = 1}^\infty G_m(t - t_0)),
			\end{align}

			where \(G(t)\) is a series of \(G_m(t)\), each of which are a function of \(H\) and its commutators. \(G_1(t)\), for example, is just \(G_1(t) = H_\mathrm{av}t\), where \(H_\mathrm{av}\) is the average value of \(H\) over the interval being integrated over, which is the solution for when \(H\) is constant in time. The terms \(G_m(t)\) for \(m > 1\) are corrections to this approximation. One can truncate the Magnus series before infinity to make a Mangus expansion, which can given an approximation for \(U(t, t_0)\). Importantly, any truncation of the Magnus expansion is unitary, meaning that probability will be conserved when used.

			Something else to note about unitary time evolution operators, is that they can be broken into a product of multiple other time evolution operators via

			\begin{align}
				U(t, t_0) &= U(t, t_1)U(t_1, t_0).
			\end{align}

			Each of these parts can be approximated as a Magnus expansion. So, if the Magnus expansion of time evolution operator is inaccurate, it can be broken down into a product of many time evolution operators which each make a much smaller step between times. These components will each have a much more accurate Magnus expansion.

			In fact, many unitary time stepping techniques have been developed based on the Magnus expansion \cite{auer_magnus_2018}. Some of these techniques use Gauss-Legendre quadrature sampling to avoid the integration of commutators of the Hamiltonian, producing simple expressions that can be used for unitary time stepping. In particular, we used the commutator free 4 (CF4) from \cite{auer_magnus_2018} as our integration method. Suppose we wish to evaluate a short time step of \(U(t_f + \mathrm{d}t, t_f)\) by taking a short step \(\mathrm{d}t\) from a fine sample time \(t_f\). Then as part of the CF4 method, the field functions are sampled at particular times based on the second order Gauss-Legendre quadrature, given by

			Then given the matrices

			\begin{align}
				\overline{G}_1 &= 2\pi\mathrm{d}t\left(\frac{3 + 2\sqrt{3}}{12}H(t_1) + \frac{3 - 2\sqrt{3}}{12}H(t_2)\right)\textrm{, and}\label{eq:cf4_sample_1}\\
				\overline{G}_2 &= 2\pi\mathrm{d}t\left(\frac{3 - 2\sqrt{3}}{12}H(t_1) + \frac{3 + 2\sqrt{3}}{12}H(t_2)\right),\label{eq:cf4_sample_2}
			\end{align}

			the unitary time step can be calculated using

			\begin{align}
				U(t_f + \mathrm{d}t, t_f) = \exp(-i\overline{G}_2)\exp(-i\overline{G}_1).\label{eq:cf4_implementation}
			\end{align}

		\subsubsection{Exponentiator}
			From Equation \eqref{eq:cf4_implementation}, evaluating matrix exponentials is a core part of the integration algorithm. However, the \texttt{spinsim} package uses the user provided field functions \(f(t)\) as a vector representation of the Hamiltonian of the spin system being simulated, rather than generic matrix elements in an array. This allows for implementation of matrix exponentiation that takes advantage of this representation. For all implementations of exponentiation, the \texttt{spinsim} computes the matrix exponential

			\begin{align}
				E(g) &= E(g_x, g_y, g_z, g_q)\\
				&= \exp(-i (g_x J_x + g_y J_y + g_z J_z + g_q Q)).
			\end{align}

			In practice, \(g\) is the Lie algebra (spin and quadrature) basis representation of \(\overline{G_1}\) or \(\overline{G_2}\) defined in Equations \eqref{eq:cf4_sample_1} and \eqref{eq:cf4_sample_2}, which need to be exponentiated in Equation \eqref{eq:cf4_implementation}. For spin half, the default exponentiator is in an analytic form. For spin one, an exponentiator based on the Lie Trotter product formula \cite{moler_nineteen_2003} is used instead. Explicitly, this formula is

			Where \(A,B\in\mathbb{C}^{N\times N}\) are matrix operators, and \(n\in\mathbb{N}\). The equality holds for without the limit for all \(n\) if the operators commute, like it does for real and complex numbers. This tells us that we can approximate a matrix exponential of a linear combination of spin operators by reducing the size of the coefficients of these operators, taking the analytically simple exponentials of each of the operators \(-iJ_x, -iJ_y, -iJ_z\) and \(-iQ\) separately, multiplying each individual result together, and then raising the combination \(T\) to the power of the original reduction.

			For our spin one case, the exponential \(E(g)\) can be approximated as, for large \(2^\tau\),

			\begin{align}
                E(g) =& \exp\left(-ig_x J_x - ig_y J_y - ig_z J_z - ig_q Q\right)\\
                =& \exp\left(2^{-\tau}\left(-ig_x J_x - ig_y J_y - ig_z J_z - ig_q Q\right)\right)^{2^\tau}\\
                \approx& \biggl(\exp\left(-i\frac12\left(2^{-\tau} g_z J_z + 2^{-\tau}g_q Q\right)\right)\nonumber\\
				&\cdot\exp\left(-i\left(2^{-\tau} g_\phi J_\phi\right)\right)\nonumber\\
				&\cdot\exp\left(-i\frac12\left(2^{-\tau} g_z J_z + 2^{-\tau} g_q Q\right)\right)\biggr)^{2^\tau}\\
                =& \begin{pmatrix}
                    \left(\frac{\Gamma}{PZ}\right)^2 & \frac{SP}{Z\Phi} & -\left(\frac{\Sigma P}{\Phi}\right)^2 \\
					\frac{SP\Phi}{Z} & CP^4 & \frac{SPZ}{\Phi} \\
					-\left(\frac{\Sigma \Phi}{P}\right)^2 & \frac{SPZ\Phi}{1} & \left(\frac{\Gamma Z}{P}\right)^2
                \end{pmatrix}^{2^\tau}\\
				=& T^{2^\tau}.
            \end{align}

            \begin{align}
				g_\phi &= \sqrt{g_x^2 + g_y^2}&&\nonumber\\
				J_\phi &= \frac{g_x}{g_\phi}J_x + \frac{g_y}{g_\phi}J_y&&\nonumber\\
				\Gamma &= \cos\left(2^{-\tau} \frac{g_\phi}{2}\right) & \Phi &= \exp\left(i2^{-\tau}g_\phi\right)\nonumber\\
				\Sigma &= \sin\left(2^{-\tau} \frac{g_\phi}{2}\right) & Z &= \exp\left(i2^{-\tau}\frac{g_z}{2}\right)\nonumber\\
				C &= \cos\left(2^{-\tau} g_\phi\right) & P &= \exp\left(i2^{-\tau}\frac{g_q}{6}\right)\nonumber\\
				S &= \frac{-i}{\sqrt2}\sin\left(2^{-\tau} g_\phi\right)&&
            \end{align}

			Note that, the methods for both spin half and spin one use analytic forms of exponentials to construct the result, meaning that all calculated time evolution operators are unitary. This guarantees that the results of \texttt{spinsim} maintain unitary and conserve probability as required.

		\subsubsection{Descretisation and parallelisation}
			Consider quantising time with
			\begin{align}
				t_k &= t_0 + \mathrm{D}t\cdot k,
			\end{align}

@@

			Beyond discretisation for parallelisation, we discretise our time evolution operator into individual integration time steps. So, each of the \(U_k\) are split into products of time evolution operators between times separated by the integration time step via

		\subsubsection{Rotating frame}
			If the rotating frame option is selected, the \(U_k\) are first calculated within a rotating frame of reference as \(U^r_k\), which, in some situations, reduces the size of the field functions used in the calculation, increasing accuracy. The rotation speed of the rotating frame is calculated locally for each parallel time step \(U_k\), and only for rotations around the \(z\) axis. The rotating from field functions \(f^r_x, f^r_y, f^r_z,\) and \(f^r_q\) are related to the field function from the user input via

@@

			Where \(f_r = f_z(t_k + \frac12\mathrm{D}t)\) is sampled the midpoint value of the fields over the interval \([t_{k - 1}, t_k]\). This, assuming that a midpoint sample is representative of an average value over the time interval, decreases the magnitude of \(f_z(t)\), while leaving the other field components at an equivalent magnitude. The rotation is then applied to obtain the lab frame time evolution operator \(U_k\) via

@@

		% \subsubsection{Magnus based integration method}
		% 	The integration method used in \texttt{spinsim} is the commutator free 4 (CF4) method from Auer et al \cite{auer_magnus_2018}, which based on the Magnus expansion. Each of the \(U_k\) are split into products of time evolution operators between times separated by a smaller time step, that is,

		% 	\begin{align}
		% 		U(t_k, t_{k-1}) &= U(t_k, t_k - \mathrm{d}t) \cdots U(t_{k-1} + 2\mathrm{d}t, t_{k-1} + \mathrm{d}t) U(t_{k-1} + \mathrm{d}t, t_{k-1})\\
		% 		U_k &= u^k_{L-1} \cdots u^k_0\textrm{, where}\\
		% 		u^k_{L-1} &= U(t_0 + (k - 1)\mathrm{D}t + (l + 1)\mathrm{d}t, t_0 + (k - 1)\mathrm{D}t + l\mathrm{d}t)
		% 	\end{align}

		% 	with \(\mathrm{d}t\) being the integration level time step. Note that the time steps are related to each other by \(\mathrm{D}t = L\mathrm{d}t\), where \(L\in\mathbb{N}\).

			% The CF4 method is used to calculate each individual \(u^k_l\). Let the fine sample time be given by \(t_f = l\mathrm{d}t + t_k\). Then as part of the CF4 method, the field functions are sampled at particular times based on the second order Gauss-Legendre quadrature, given by

			% \begin{align}
			% 	t_1 &= t_f + \frac12 \mathrm{d}t\left(1 - \frac{1}{\sqrt{3}}\right)\textrm{, and}\\
			% 	t_2 &= t_f + \frac12 \mathrm{d}t\left(1 + \frac{1}{\sqrt{3}}\right).
			% \end{align}

			% Now, let
			% \begin{align}
			% 	f(t_1) &= (f_x(t_1), f_y(t_1), f_z(t_1), f_q(t_1))\textrm{, and}
			% 	f(t_2) &= (f_x(t_2), f_y(t_2), f_z(t_2), f_q(t_2)).
			% \end{align}

			% The integration time evolution operator can then be calculated using

			% \begin{align}
			% 	g_1 =& (g_{1,x}, g_{1,y}, g_{1,z}, g_{1,q})\\
			% 	=& 2 \pi \mathrm{d}t \left(\frac{3 + 2 \sqrt{3}}{12} f(t_1) + \frac{3 - 2 \sqrt{3}}{12} f(t_2)\right)\textrm{, and}\\
			% 	g_2 =& (g_{2,x}, g_{2,y}, g_{2,z}, g_{2,q})\\
			% 	=& 2 \pi \mathrm{d}t \left(\frac{3 - 2 \sqrt{3}}{12} f(t_1) + \frac{3 + 2 \sqrt{3}}{12} f(t_2)\right)\textrm{, so}\\
			% 	u =& \exp(-i \left( g_{2,x} J_x + g_{2,y} J_y + g_{2,z} J_z + g_{2,q} Q\right))\\
			% 	&\cdot\exp(-i \left( g_{1,x} J_x + g_{1,y} J_y + g_{1,z} J_z + g_{1,q} Q\right)).
			% \end{align}

	\subsection{Software architecture}
		\subsubsection{Integrator architecture}
			The integrator in the \texttt{spinsim} package calls a \texttt{numba.cuda.jit()}ed kernel to be run on a \emph{Cuda} capable \emph{Nvidia} GPU in parallel, with a different thread being allocated to each of the \(U_k\). This returns when each of the \(U_k\) have been evaluated.

			The thread starts by calculating \(t_k\) and, if the rotating frame is being used, \(f_r\). The latter is done by sampling a (\texttt{numba.cuda.jit()}ed version of a) user provided \emph{python} function \(f\) describing how to sample the Hamiltonian. The code then loops over each integration time step \(\mathrm{d}t\) to calculate the integration time evolution operators \(u^k_l\).

@@

			where \(j \in \{\frac12, 1\}\) is the spin quantum number of the system.
